Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-09-09 alloscore_manuscript/manuscript.tex
fix 'highest at at'
response2, using hyperlinks to diff

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Aug 20 15:20:10 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
\texttt{CU-select} & 669 & 133 & 774 & 326 \\ 
  \texttt{per-capita} & 865 & - & 1029 & 366 \\ 
  \texttt{COVIDhub-ensemble} & 873 & 159 & 1067 & 438 \\ 
  \texttt{USC-SI\_kJalpha} & 995 & 91 & 1216 & 1097 \\ 
  \texttt{JHUAPL-Gecko} & 1034 & 164 & 1141 & 418 \\ 
  \texttt{MUNI-ARIMA} & 1084 & 169 & 1248 & 440 \\ 
  \texttt{COVIDhub-trained\_ensemble} & 1089 & 169 & 1271 & 823 \\ 
  \texttt{COVIDhub-baseline} & 1175 & 170 & 1317 & 535 \\ 
  \texttt{JHUAPL-Bucky} & 1358 & 102 & 1566 & 1214 \\ 
  \texttt{JHUAPL-SLPHospEns} & 1540 & 129 & 1604 & 1102 \\ 
  \texttt{UVA-Ensemble} & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \ref{sec:ias_examp}). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Tue Aug 20 15:20:15 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
model & MAS & taMWIS & taMWIS rank \\ 
  \hline
\texttt{COVIDhub-ensemble} & 389 & 70 & 2 \\ 
  \texttt{per-capita} & 464 & - & - \\ 
  \texttt{COVIDhub-trained\_ensemble} & 483 & 87 & 4 \\ 
  \texttt{CU-select} & 502 & 81 & 3 \\ 
  \texttt{JHUAPL-SLPHospEns} & 526 & 67 & 1 \\ 
  \texttt{COVIDhub-baseline} & 594 & 93 & 5 \\ 
  \texttt{JHUAPL-Bucky} & 643 & 112 & 7 \\ 
  \texttt{MUNI-ARIMA} & 707 & 116 & 8 \\ 
  \texttt{JHUAPL-Gecko} & 929 & 110 & 6 \\ 
  \texttt{USC-SI\_kJalpha} & 1473 & 155 & 9 \\ 
   \hline
\end{tabular}
\caption{Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the \texttt{per-capita} allocation which does not use forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

@@

AS was computed for a range of $K$ from 200 to 60,000 at increments of 200 for forecasts made on December 20, 2020 predicting levels of
hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These
calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
19,581 that day. Model ranks by AS were fairly stable across all $K$, and there appears
to be an substantial interval around the observed value over which ranks are constant.

@@

Probabilistic forecasts of infectious disease outbreaks have typically been evaluated using well-known proper scoring
rules such as the LS, CRPS, and WIS. Often models are ranked according to such a score, but without reference to any
underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated. 
We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by
including evaluations of forecasts with
scoring rules that directly consider the success of forecasts in supporting specific decisions. The allocation score was developed for this purpose. The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecast scores. The benefits of the approach to forecast evaluation discussed here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
can be clearly juxtaposed with standard accuracy measures of the forecast \citep{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \citep{syme1999defining, gebre2021multi, colett2016using}.

>>>>

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Mon Sep  9 16:35:24 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
\texttt{CU-select} & 669 & 133 & 774 & 326 \\ 
  \texttt{per-capita} & 865 & - & 1029 & 366 \\ 
  \texttt{COVIDhub-ensemble} & 873 & 159 & 1067 & 438 \\ 
  \texttt{USC-SI\_kJalpha} & 995 & 91 & 1216 & 1097 \\ 
  \texttt{JHUAPL-Gecko} & 1034 & 164 & 1141 & 418 \\ 
  \texttt{MUNI-ARIMA} & 1084 & 169 & 1248 & 440 \\ 
  \texttt{COVIDhub-trained\_ensemble} & 1089 & 169 & 1271 & 823 \\ 
  \texttt{COVIDhub-baseline} & 1175 & 170 & 1317 & 535 \\ 
  \texttt{JHUAPL-Bucky} & 1358 & 102 & 1566 & 1214 \\ 
  \texttt{JHUAPL-SLPHospEns} & 1540 & 129 & 1604 & 1102 \\ 
  \texttt{UVA-Ensemble} & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS), (see Section \ref{sec:ias_examp}). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

% latex table generated in R 4.4.1 by xtable 1.8-4 package
% Mon Sep  9 16:35:29 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrr}
  \hline
model & MAS & taMWIS & taMWIS rank \\ 
  \hline
\texttt{COVIDhub-ensemble} & 389 & 70 & 2 \\ 
  \texttt{per-capita} & 464 & - & - \\ 
  \texttt{COVIDhub-trained\_ensemble} & 483 & 87 & 4 \\ 
  \texttt{CU-select} & 502 & 81 & 3 \\ 
  \texttt{JHUAPL-SLPHospEns} & 526 & 67 & 1 \\ 
  \texttt{COVIDhub-baseline} & 594 & 93 & 5 \\ 
  \texttt{JHUAPL-Bucky} & 643 & 112 & 7 \\ 
  \texttt{MUNI-ARIMA} & 707 & 116 & 8 \\ 
  \texttt{JHUAPL-Gecko} & 929 & 110 & 6 \\ 
  \texttt{USC-SI\_kJalpha} & 1473 & 155 & 9 \\ 
   \hline
\end{tabular}
\caption{Time averaged mean weighted interval scores (taMWIS) and mean allocation scores (MAS) by model across 13 weeks. These results show the average performance across time for the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, as well
  as for a per-capita allocation. MWIS is not defined for the \texttt{per-capita} allocation which does not use forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

@@

AS was computed for a range of $K$ from 200 to 60,000 at increments of 200 for forecasts made on December 20, 2020 predicting levels of
hospitalizations on January 3, 2021 as well as for the \texttt{per-capita} allocation (Figure \ref{fig:multi-k}A). These
calculations highlight that AS was highest at at values of $K$ near the observed nationwide total hospital admissions of
19,581 that day. Model ranks by AS were fairly stable across all $K$, and there appears
to be a substantial interval around the observed value over which ranks are constant.

@@

Probabilistic forecasts of infectious disease outbreaks have typically been evaluated using well-known proper scoring
rules such as the LS, CRPS, and WIS. Often models are ranked according to such a score, but without reference to any
underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated. 
We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by
including evaluations of forecasts with
scoring rules that directly consider the success of forecasts in supporting specific decisions. The allocation score was developed for this purpose. The AS is expressed in units that are directly relevant to decisions, which helps not only with ranking forecasts in the context of a specific action, but also allows decision-makers to contextualize the magnitude of differences between forecast scores.  For example, it may be the case that differences in AS are small relative to the amount of available resources or the accuracy with which planned resource allocations 
can be executed.
The benefits of the approach to forecast evaluation discussed here generalize beyond public health, to any application of probability forecasts for decision-making, such as economics, where the monetary consequences of using a forecast
can be clearly juxtaposed with standard accuracy measures of the forecast \citep{bannigidadmath2016stock, zhang2019forecasting}, resource allocation in urban planning \citep{burnett2014urban, liang2013resource}, or sustainability initiatives such as water and energy allocation \citep{syme1999defining, gebre2021multi, colett2016using}.
